# Day 2

## Outline for Day 2

Disclaimer:
The focus of the training is not on understanding the statistical concepts, but rather on how to implement common tasks in R - with a few stats sprinkles.
I'm trying to show you the quick & basic way of doing this and give you an idea of the fancier approaches - which gives you more flexibility (but also has a steeper learning curve).
You can of course dive deeper on those topics or request follow-up training. :)


Today we will cover the following topics:

* plotting (base R, ggplot2)
* regression models
* testing regression assumptions
* significance tests (chi2, t-test)

### Take home messages to start with 

1. Work on a copy of your data in R - do not touch the raw data
2. Work in R projects 
3. Structure your code with sections (Ctr shift R)
4. help(x) or "?x" is your friend when trying to understand function x... scroll down to "Examples" in the helper window that pops up

# Loading packages and data
```{r}
# load all required packages
pacman::p_load(dplyr, readr, ggplot2, corrplot, NHANES, survey, broom, lmtest, car, knitr)
```

```{r, echo=FALSE, message= FALSE}
# loading it all in one go as above is very neat - here you can see how you'd 
# usually call individual packages and what each is used for in this course:

library(dplyr) # for useful functions and code syntax (e.g. %>% pipes)
library(readr) # for easy reading in of data and saving data in formats such as csv (also tsv or rds)
library(ggplot2) # for more intermediate / advanced plotting
library(corrplot) # for plotting a correlation matrix very easily
library(NHANES) # for easily getting the data
library(survey) # for using survey weights and using versions of functions that allow for survey weights
library(broom) # for tidying up model outputs
library(lmtest) # for heteroscedasticity tests
library(car) # for variance inflation factor function (for multicollinearity)
library(knitr) # not relevant for this course, but this helps displaying the output from this script in the html format that you can then see in your browser
```


```{r, echo=FALSE}
# load data
nhanes <- NHANES::NHANESraw

```


```{r}
# inspect data by calling head() function
# head(nhanes)
knitr::kable(head(nhanes),format = 'html')

```

```{r}
dplyr::glimpse(nhanes) # if there are many columns (from dplyr package)
```

```{r}
# get a feeling for variables
# categorical
levels(nhanes$Depressed)
table(nhanes$Depressed,  useNA = "ifany")
# numerical
summary(nhanes$BMI, useNA = "ifany")

```
