
# Plotting basic charts (base R) 

It is important to distinguish between exploratory graphs and explanatory graphs:
* Exploratory is done as part of analysis and there is no need to be pretty
* Explanatory graphs are done once we understand the data and want to get insights across (for sharing with others)

## histograms - basic frequencies
```{r}
# basic frequencies for numerical variables
hist(nhanes$Height)
hist(nhanes$Age) 
hist(nhanes$Age[nhanes$Age<10]) # they sampled for both 0-6months and 6-12months

```

## bar charts - plotting stats across categories
```{r}

# Great for plotting a statistic (e.g. mean value) for categorical data
# Get dataframe into right format
aggregates_by_health <- nhanes %>% 
  dplyr::group_by(HealthGen) %>% 
  dplyr::summarise(count = n(),
            mean_weight = round(mean(Weight, na.rm=T),1),
            mean_age = round(mean(Age, na.rm=T),1))

# create bar plot to plot either Frequencies or aggregate statistics
# Participant numbers by race
barplot(count ~ HealthGen, data = aggregates_by_health)
# average weight and age by race
barplot(mean_weight ~ HealthGen, data = aggregates_by_health)


```

## Box plots - plotting distribution of several categories/vars
```{r}

# Good for distribution of continuous data across categories
boxplot(nhanes$Age ~ HealthGen, data = nhanes )

```

## Scatter plots - relationship between two continuous vars 
```{r}

# Let's keep only adults in for this
adults <- nhanes %>% dplyr::filter(Age >= 18)

# plot scatter plot
plot(x = adults$Height, y = adults$Weight)

```

Can you guess the Correlation coefficient?
What's the value of the correlation coefficient?"
```{r}
cor(adults$Height, adults$Weight, use = "pairwise.complete.obs")
```

Now let's add the line of best fit.
First, we calculate slope and intercept of line of best fit

```{r}
coef(lm(Weight ~ Height, data = adults))
```


Then we can add them to the plot
```{r}
# add them to plot (run both commands)
plot(x = adults$Height, y = adults$Weight)
# intercept and slope
abline(-72.209450 , 0.915329)
```

## Tiny statistics excursion
What's the relationship between the linear model regression coefficient and the correlation coefficient?


```{r}

hnorm <- adults$Height/sd(adults$Height, na.rm = TRUE)
wnorm <- adults$Weight/sd(adults$Weight, na.rm = TRUE)
df_norm <- as.data.frame(cbind(hnorm,wnorm))

coef(lm(hnorm ~ wnorm, data = df_norm)) 

```
Note: For correlation coef, the larger the value, the stronger the (linear!) relationship.
For regression coefficients, a larger slope coefficient does NOT imply that the association between variables is stronger (remember the unit dependency).

## Are there any other plots that you regularly use?
