# Intermediate plotting in R (ggplot2)

This first bit of plotting was hopefully all pretty straight forward and the code was easy to read? In the R universe new packages are created all the time and ggplot2 is *THE* data viz package in R.  

It allows for more customisable ways of plotting -  we can add layer, upon layer of extra info and change transparency and colours and much, much more. In a nutshell we have:

1. DATA, 
2. AESTHETICS, 
3. GEOMETRIES.

If you are interested in finding out more about how graphics are constructed with these three components, have a search around your favourite platform for "Grammar of Graphics" and check out the documentation of the ggplot2 package (https://ggplot2.tidyverse.org/).   
Let's have a look at what this means in practice.

## Adding transparency

```{r, message=FALSE}
# 2.1 Transparency
ggplot(data = adults, aes(x = Height, y = Weight)) +
  geom_point(alpha = 0.4)

# explore the warning message
# sum(is.na(adults$Weight)|is.na(adults$Height))
```

## Adding automatic line of best fit

```{r, message=FALSE, warning= FALSE}
# 2.2 scatter plot with automatic line of best fit
ggplot(data = adults, aes(x = Height, y = Weight)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) # se = TRUE would show standard error bars along the line
```

## Adding colours

Other cool things you can easily / automatically do with ggplot include colouring by category:

```{r}

# Define default colour scale suitable for colour-blind users

scale_colour_discrete <- ggthemes::scale_color_colorblind 
```

```{r, message=FALSE, warning= FALSE}

# plot in different colours based on the Gender variable

ggplot(data = adults, aes(x = Height, y = Weight, colour = Gender)) +
  geom_point(alpha = 0.4)

```

## Fitting a line of best fit for each group of a categorical variable

```{r, message=FALSE, warning= FALSE}

# 2.4 get best fit line by Species
ggplot(data = adults, aes(x = Height, y = Weight, color = Gender)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE)

```

## Exercise: Brief in-class practice of making charts 

Try out the simple commands for the charts that you most frequently produce in SPSS / SAS.

Then choose a variable of interest that is continuous (e.g. Height or Pulse)  and produce a box plot showing a break-down of distribution by a categorical variable of interest (e.g. HealthGen, MaritalStatus).

## ggplot cheat sheet

If you're lost or want to explore more options for plotting charts, you can consult a ggplot cheat sheet, e.g. https://www.rstudio.com/wp-content/uploads/2016/11/ggplot2-cheatsheet-2.1.pdf

