# Intermediate plotting in R (GGPLOT2)

This was hopefully all very straight forward and the code was easy to read. In the R universe new packages are created all the time
GGPLOT2 is THE data viz package in R
More customisable ways of plotting - in R people use a package called 'ggplot2' we can add layer, upon layer of extra info and change transparency and colours in a nutshell we have 

1. DATA, 
2. AESTHETICS, 
3. GEOMETRIES.

Let's have a look at what this means in practice.

## Adding transparency

```{r, message=FALSE}
# 2.1 Transparency
ggplot(data = adults, aes(x = Height, y = Weight)) +
  geom_point(alpha = 0.4)

```

## Adding automatic line of best fit

```{r, message=FALSE, warning= FALSE}
# 2.2 scatter plot with automatic line of best fit
ggplot(data = adults, aes(x = Height, y = Weight)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) # se = TRUE would show standard error bars along the line
```

## Adding colours

Other cool things you can easily / automatically do with ggplot include colouring by category:

```{r}

# Define default colour scale suitable for colour-blind users

scale_colour_discrete <- ggthemes::scale_color_colorblind 
```

```{r, message=FALSE, warning= FALSE}

# plot in different colours based on the Gender variable

ggplot(data = adults, aes(x = Height, y = Weight, colour = Gender)) +
  geom_point(alpha = 0.4)

```

## Fitting a line of best fit for each group of a categorical variable

```{r, message=FALSE, warning= FALSE}

# 2.4 get best fit line by Species
ggplot(data = adults, aes(x = Height, y = Weight, color = Gender)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE)

```

## Exercise: Brief in-class practice of making charts 

Try out the simple commands for the charts that you most frequently produce in SPSS / SAS.

Then choose a variable of interest that is continuous (e.g. Height or Pulse)  and produce a box plot showing a break-down of distribution by a categorical variable of interest (e.g. HealthGen, MaritalStatus).

## ggplot cheat sheet

If you're lost or want to explore more options for plotting charts, you can consult a ggplot cheat sheet, e.g. https://www.rstudio.com/wp-content/uploads/2016/11/ggplot2-cheatsheet-2.1.pdf

