
# Significance tests

## Chi-square 
Chi-square test for categorical variables determines whether there is a difference in the population proportions between two or more groups

Let's look at smoking for men vs. women

```{r}
contingency_table_gender <- table(adults$Gender,
                           adults$SmokeNow)

contingency_table_gender

```

Or the 'dplyr way':

```{r}
adults %>% 
  dplyr::filter(!is.na(SmokeNow)) %>% 
  dplyr::group_by(Gender, SmokeNow) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::mutate(freq = n / sum(n))
```

```{r}
chisq.test(adults$SmokeNow, adults$Gender)
```

## T-test 

You can specify whether you need it to be one-sided / two-sided & one-sample / independent-sample.

Let's look at whether BMI differs between men and women:

```{r}

hist(adults$BMI[adults$Gender=="female"])
hist(adults$BMI[adults$Gender=="male"])

```
Or let's do it the ggplot2 way


```{r}
# Define default colour scale suitable for colour-blind users

scale_colour_discrete <- ggthemes::scale_color_colorblind
```

```{r, message=FALSE}
# Create plot facetted by Gender
ggplot(adults, aes(BMI, fill = Gender)) +
  geom_histogram() +
  facet_wrap(.~Gender)
```

Any guesses on whether there is a significant difference?
Now it's time to do the test.
```{r}
t.test(x = adults$BMI[adults$Gender == "female"],
       y = adults$BMI[adults$Gender == "male"],
       alternative = "two.sided", # could also be 'less' or 'greater' for one-sided test
       paired = FALSE) 
```
Yes, on average women seem to have a higher BMI in this survey.
