
# 'Homework' questions

The code for sample solutions is given in the file Homework_solutions_day2 file.
Please try to answer the following questions:

1. Has the relationship between height and weight changed between the survey years?

2. Is there a significant difference in how many rooms individuals' accommodation has, between those who own and those who rent their accommodation?

3. How are trouble sleeping and depression (both categorical) related?

## Training evaluation

Please fill in this survey before you leave the call:

https://forms.gle/J4Yt64C1FAncFFjW6




# Extra: Survey specific functions

As James showed you yesterday if we are using survey data that comes with weights, we need to account for those.

The 'survey' package has adapted functions for the different things that we did today.
Let's first create the weights for the data at hand, specify the sample design and store it in an R object that we can then call in our tests.

```{r survey weights}
# add weighting variable
nhanes <- nhanes %>% dplyr::mutate(WTMEC4YR = 0.5 * WTMEC2YR)

# Create weighted R data object that holds the information about the sampling specifications
nhanes_design <- survey::svydesign(data = nhanes, 
                              strata = ~SDMVSTRA, 
                              id = ~SDMVPSU, 
                              nest = TRUE, 
                              weights = ~WTMEC4YR)

```

Now we can seemlessly work with the weighted data. Let's create a frequency table that accounts for sample weights.
```{r survey table}

table_gender_smoke_w <- survey::svytable(~Gender + SmokeNow, design = nhanes_design) # contains estimated frequencies based on survey weights

table_gender_smoke_w

```

For reference, there is some sample code for the other main analyses using the survey package.


#### Survey-weighted histogram

```{r}
# BMI
ggplot(data = nhanes,
       mapping = aes(BMI, weight = WTMEC4YR)) + 
	geom_histogram(binwidth = 2,
                   color = "white") +
	labs(x = "BMI")

```

#### Survey-weighted bar plot 

Bar plot of SleepHrsNight by gender

```{r}

# Compute the survey-weighted mean by Gender
Sleep_mean_gender <- survey::svyby(formula = ~SleepHrsNight, 
             by = ~Gender, 
             design = nhanes_design, 
             FUN = svymean, 
             na.rm = TRUE, 
             keep.names = FALSE)
             
# Construct a bar plot of average SleepHrsNight by gender
ggplot(data = Sleep_mean_gender, mapping = aes(x = Gender, y = SleepHrsNight)) +
  geom_col() + 
  labs("Average Hours of Sleep")

```

#### Survey-weighted Chi2 test

Next, let's conduct a Chi2 test with survey weights

```{r survey Chisq}
survey::svychisq(~Gender + SmokeNow, 
         design = nhanes_design,
         statistic = "Chisq")
```

#### Survey-weighted t-test

```{r}
# do mean sleeping hours differ between men and women
svyttest(formula = SleepHrsNight~Gender,
       design = nhanes_design)
```
```{r}
# Let's get the mean values printed out (created in step prior to t-test) to aid interpretation of the t-test
Sleep_mean_gender
```
#### Survey-weighted scatter plot

Let's look at the relationship between height and weight in 8 year old children.
You can set the transparency parameter alpha to be the sample weight (WTMEC4YR). Observations with a bigger weight will show as darker points.

```{r}
ggplot(data = nhanes %>% filter(Age == 8),
       mapping = aes(x = Height, y = Weight, alpha = WTMEC4YR)) + 
    geom_point() + 
    guides(alpha = FALSE)
```

#### Survey-weighted regression

Let's model Weight based on Height for 8 year olds
```{r}
model <- svyglm(Weight ~ Height, data = nhanes %>% filter(Age == 8),
                design = nhanes_design)

summary(model)
```
